\section{General Background of Set Theory \& Mathematical Logic\\Mathematical Logic (i.e. Symbolic Logic)}

Make it clear that there is no generally accepted definition of logic.  But there is a definition of a syllogism.  This is sort of the same thing as saying there is no definition of \emph{poetry} but a \emph{hiaku} is very well defined.


\ifKey
\color{red}
The followig is an example from the notes about a syllogism.  
\begin{itemize}
\item {\bf Major premise:} Sixty men can do a piece of work Sixty times as quickly as one man.
\item {\bf Minor premise:} One man can dig a post hole in Sixty seconds.
\item {\bf Conclusion:} Therefore Sixty men can take a post hole in one second.
\end{itemize}

Perhaps you can discuss the obvious lack of validity to this argument.  Infact this is a nice way to begin the discussion about sound and valid arguments.  

\color{black}
\else
\notes{5}
\fi

\ifKey
\color{red}
Clarify to students thatt syllogisms have two basic properties: validity and soundness. A syllogism is \emph{valid} if the conclusion is a logical consequence of the premises.  A \emph{sound} syllogism is one that is not only valid, but the premises are also true.
\color{black}
\else
\notes{5}
\fi



The following problems would be great for break-out discussions among groups of students.  
\begin{problem} Complete this syllogism, and determine if the syllogism valid and sound?

\ifKey
\color{red}
\begin{enumerate}
\item [$-$]All fish can breath underwater.
\item [$-$]Socrates was a fish.
\item [$\star$]\notes{1}
\end{enumerate}
\vspace{1em}
\color{black}
\else
\notes{5}
\fi

\end{problem}

The following two are a little harder.  Both of these arugments are valid.
\begin{problem}Consider this syllogism.

\ifKey
\color{red}
\begin{enumerate}
\item [$-$] Everybody loves my baby.
\item [$-$] My baby loves only me.
\item [$\star$] Therefore, I am my own baby.
\end{enumerate}

Explain why  the syllogism valid.

\vspace{1em}

\color{black}
\else
\notes{5}
\fi
\end{problem}


\begin{problem}Consider this syllogism.

\ifKey
\color{red}
\begin{enumerate}
\item [$-$] Everyone loves a lover.
\item [$-$] Romeo loves Juliet.
\item [$\star$] Therefore Trump loves Obama.
\end{enumerate}

Explain why  the syllogism valid.

\vspace{1em}

\color{black}
\else
\notes{5}
\fi
\end{problem}


