\section{Sets and Vampires}

\begin{problem}[Two Vampires, Vladamir and  Countess Elizabeth Bathory] 
Pretend two Vampires, Vladamir and  Countess Elizabeth Bathory, are playing a guessing game.  The Countess writes a positive integer on the inside of her coffin.  Each day Vladamir gets one guess at the number.  What would be a strategy for Vladamir to eventually guess the correct number the Countess has written on the inside of her coffin?
\notes{8}
\end{problem}
\begin{problem}[Vampires are playing another guessing game.]
Now the Countess writes \emph{any} integer on the inside of her coffin.  Again each day Vladamir gets one guess at the number, and if he is correct, he wins.  What would be a strategy for Vladamir to eventuallyguess the correct number the Countess has written on the inside of her coffin?
\notes{8}
\end{problem}

\begin{problem}[Let's increase the difficulty.]
Now the Countess writes a pair of positive integers on the inside of her coffin.  Again each day Vladamir gets one guess at the two numbers; he has to get both correct on the same day.  What would be a strategy for Vladamir to eventually guess the pair of numbers the Countess has written on the inside of her coffin?
\notes{8}
\end{problem}

\newpage
\begin{problem}[The Countess writes a fraction on the inside of her coffin]
Ok, let the Countess write a fraction on the inside of her coffin.  Again each day Vladamir gets one guess at the fraction.  What would be a strategy for Vladamir to win?
\notes{8}
\end{problem}

\begin{problem}[Now the Countess writes a finite set of positive integers.]
This is it!   Now the Countess writes a finite set of positive integers.  Again each day Vladamir gets one guess at the set.  How can Vlademir win? 
\notes{8}\label{vampire5}
\end{problem}

What Problem \ref{vampire5} is really saying is that the set of all finite sets of positive integers is denumerable.  But here is the kicker.  The set of all sets of positive integers (both finite and infinite) is NOT denumerable!  This discovery by Georg Cantor marked the beginning of Set Theory in 1878, Althought his paper did not include vampires.

\newpage
\begin{problem}[Now the Countess writes an infintie set of positive integers.]
It seems like Vladamir can always win (and this is originally what Cantor thought too).  But now assume that the Countess writes or describes a  set of positive integers on the inside of her coffin, but this time it can be an infinite set.  Again each day Vladamir gets one guess at the set.  

However Vladamir has a secret weapon, a book with an infinite number of pages, page-1, page-2, ...page-k,...  On each page is written a set (or a description of a set since infinite sets are hard to write on a page).  He is going to use this to eventually guess the Countess' set, day by day, page by page.  

Unbeknownst to Vladamir, the Countess found a copy of Vladamir's  book on Amazon.  Is there a way that the Countess can describe a set that is not in the book?
\notes{10}\label{vampire6}
\end{problem}

